Type-Safe Modular Parsing

### 2017-01-14 paper/sections/Conclusion.tex
Revised 4.4

Abstracting language features as reuseable components can be achieved based on our solution, which is useful for rapid language prototyping. Our solution does not rely on particular techniques. On the contrary, it is a general framework customizable by the user. The TAPL case study shows that we can obtain considerable code reuse (69\%) than a non-modular implementation, using our modular parsing approach and language feature abstraction.

>>>>

Abstracting language features as reuseable components can be achieved based on our solution, which is useful for rapid language prototyping. Our solution does not rely on particular techniques. On the contrary, it is a general framework customizable by the user. The TAPL case study shows that we can obtain considerable code reuse (69\% shorter) comparing with a non-modular implementation, using our modular parsing approach and language feature abstraction.

### 2017-01-14 paper/sections/ObjectAlgebras.tex
Revised 4.4

\subsection{Composition of Algebras}\label{subsec:more-oa}
\huang{to revise}
By adopting Object Algebras, we also inherit some advanced features from the pattern. Specifically, in the original paper~\cite{Oliveira:2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}. A query traverses the abstract syntax tree, collects information and returns the result. The aforementioned \lstinline{Print} and \lstinline{Eval} are traditional queries. But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing. It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice. Instead, we can firstly compose two queries into one, then feed it to the parser. For space reasons we omit the code here.

Another interesting observation, if we look back at the end of Section~\ref{subsec:objectalgebras}, is that \lstinline{makeExp} represents an ``object'' of the expression AST. Or rather, what OA parsers generate are ``implicit objects'', which are actually functions of type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}. Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them. Especially, in the design of a compiler
there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
in a very clever way. Continuing from the old example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}

\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result. Furthermore, transformations can be composed in a linear pipeline-style, before a query algebra is fed ~\cite{Zhang2015}.

>>>>

\subsection{Discussion of Object Algebras}\label{subsec:more-oa}

In Section~\ref{subsec:objectalgebras}, \lstinline{makeExp} represents the AST of expression \inlinecode{1 + x} as a method. Parsers with Object Algebras generate such ``implicit objects'', which are actually functions of type \lstinline{Alg[E] => } \lstinline{E} for generic \lstinline{E}. Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them. Especially, in the design of a compiler there will be a lot of desugarings (or AST transformations).

However, we argue that this is not an issue, because functions can be used
in a very clever way. Continuing with the previous example, suppose we want to transform an expression of \inlinecode{VarExprAlg} to an expression of \inlinecode{ExprAlg}, by
replacing some variables with values based on a variable environment. A \textit{transformation} algebra can be implemented as follows:

\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
Here the \inlinecode{VarExprAlg} expression \inlinecode{1 + x} is changed to \inlinecode{ExprAlg} expression \inlinecode{1 + 2}, before applying the pretty-printing algebra. The abstract field
\inlinecode{alg} in \inlinecode{Refactor} could be any algebra simply for delegation. With an algebra instance of \inlinecode{alg}, \inlinecode{Refactor} is able to transform a \inlinecode{VarExprAlg} expression before producing the result. Furthermore, transformations can be composed in a linear pipeline-style beforehand~\cite{Zhang2015}.

Another useful technique is the composition of \textit{query} algebras. A query traverses the abstract syntax tree, collects information and returns the result. The aforementioned \lstinline{Print} and \lstinline{Eval} algebras are traditional queries. But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing. It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice. Instead, we can firstly compose two queries into one~\cite{Oliveira:2012,oliveira2013feature}, before applying it for parsing.

### 2017-01-14 paper/sections/PackratParsing.tex
Revised 4.4

\noindent However, this results in the worst case exponential time
complexity in Parsec, because it does not have proive optimizations
for backtracking.

>>>>

\noindent However, this results in the worst case exponential time
complexity in Parsec, because it does not provide optimizations
for backtracking.
